The effect of clumpy outflows and disc anisotropy on quasar unification scenarios

### 2015-07-15 draft.tex
a bunch of text in results, may clash with laptop

\author[Matthews et al.]{J.~H.~Matthews$^1$\thanks{jm8g08@soton.ac.uk}, C.~Knigge,$^1$
N.~Higginbottom,$^1$ K.~S.~Long,$^2$ S.~A.~Sim,$^3$ and S.~W.~Mangham
\medskip  
\\$^1$School of Physics and Astronomy, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom
\\$^2$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD, 21218
\\$^3$School of Mathematics and Physics, Queens University Belfast, University Road, Belfast, BT7 1NN, Northern Ireland, UK}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%          INTRODUCTION
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Introduction focussing on key points

\begin{itemize}
\item standard wind + BALQSO introduction
\item focus on unification and that we are testing it
\item some discussion of scales, referencing e.g. reverb maps, variability, microlensing, Arav 
\item clumping background: stellar winds, clumping in AGN winds, variability
\end{itemize}
%%\item

% Outflows are ubiquitous in active galactic nuclei (AGN) and quasi-stellar objects (QSOs).
% They can take the form of highly collimated radio jets \citep{bellonijet2010}, 
% or mass-loaded `winds' emanating from the accretion disk. 
Approximately $20\%$ of QSOs exhibit broad absorption lines (BALs) in the ultraviolet,
providing clear evidence for outflowing absorbing material
\citep{weymann1991, knigge2008, turnermiller2009,allen2011}.
The simplest explanation for the incidence of 
BALs in quasar samples is in terms of an accretion disk wind (ADW). 
In this paradigm, the BALQSO fraction is associated with
the covering factor of the outflow.
Indeed, many have gone further, and suggested that the diverse phenemonology
of luminous AGN and QSOs can be broadly explained by
into a simple picture of geometric unification \citep[e.g.]{MCGV95, elvis2000}. 
In such a model, a biconical wind rises from 
a thin accretion disk \cite{shakurasunyaev1973}, approaching velocities
close to $\sim0.1c$. Depending on viewing angle, an observer 
may then see a BALQSO, LoBAL-QSO or normal `Type 1' quasar- with 
broad emission lines superposed on a disk-like continuum.

The UV spectrum of {\em non-BAL} QSOs is typified by a series
of strong emission lines (e.g. \la, \civ, \nv) with an underlying blue continuum
- the so-called {\sl `big blue bump'} (BBB). The BBB is normally attributed to blackbody-like
emission from an accretion disk surrounding the central black hole (REF), similar
to that described by \cite{shakurasunyaev1973}. However,
a number of issues have arisen relating to this model. First, 
AGN/QSO spectra exhibit a `break' in the spectrum at around $1000$~\AA 
which scales only weakly with black hole mass. This potentially 
suggests a problem with a thin disk model (e.g. Antonucci). 
However, it is possible that this is the result of incorrect intergalactic medium
(IGM) corrections (REF) or corresponds to the temperature 
at which a line-driven wind carries mass away from the system (laor davis).
Second, results from microlensing (REFs) imply that the disk emission 
region is $\sim4$ times larger than one might expect from a Shakura-Sunyaev
model. Inhomogenous disks have been proposed as an alternative (REFs).
These observations appear to pose problems for the thin disk model of luminous,
sub-Eddington AGN. However, recent results from \cite{capellupo2015} find 
that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
then AGN SEDs can, in general, be fitted well with accretion disk models.
Uncovering the intrinsic disk SED and understanding the effect of the outflow on the accretion 
mechanism is clearly crucial if we are to properly understand the physical nature
of AGN.

The geometry and size of the BLR is also a matter of contention. 
The main constraints on the emission region size come from microlensing (REFs)
and reverberation mapping (REFs). While these observations have mostly been
carried out for Seyfert galaxies, there are also a few instances of these methods
being applied to quasars (REFs). A number of different proposals have been made 
for the origin of the BLR. Early AGN unification scenarios posited that the broad emission lines
where produced by clouds of plasma orbiting fairly close to the BH (refs).
Since then, multiple interpretations of a disk wind model have been proposed,
with varying radii and geometries (refs).

In this paper, we attempt to test the disk wind paradigm 
using Monte Carlo radiative transfer and photoionization calculations.
In section 2, we describe our code. In section 3, we briefly discuss some 
of the successes and failures of our previous benchmark model for BALQSOs 
\citep[][hereafter H13]{higginbottom2013} and outline the model, including 
a description of our clumping prescription. In section 4, we present the results 
from a clumped model which successfully reproduces
the observed ionization state while maintaing realistic X-ray properties.
In section 5 we discuss our results, focussing particular on the anisotropy of 
disk emission and GR effects, and finally, in section 6, we summarise our findings.

@@

\section{Method: Ionization and Radiative Transfer}

We use a code Python described extensively by LK02, S05, H13, M15...

@@

We adopt the same hybrid scheme described by \cite{M15}...
% , but we have improved our ionization scheme
% for simple ions to explicitly solve the rate equations based on MC estimators.

\subsection{Macro-atoms}

\section{A Simple Biconical Disk Wind Model for QSOs}

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/fig2_cartoon.png}
\caption
{
A cartoon showing the geometry and some key parameters of
our biconical wind model.
}
\label{fig:alpha_ox}
\end{figure}

\subsection{A Benchmark Model for BALQSOs}

\cite{higginbottom2013} presented a benchmark model for
(BAL)QSOs...introduce key parameters. 
% which successfully reproduced BAL features 
% in synthetic spectra using \py. However, the model
% had a few key drawbacks when considering its potential
% for unification. First, it failed to produce significant
% broad emission lines at non-BAL viewing angles; this is 
% a key requirement for a unified QSO model. Second,
% the model was significantly X-ray weak, as it was found
% that an X-ray luminosity comparable to that of a QSO 
% over-ionized the wind, resulting in no UV resonance 
% line absorption features.

% \begin{table}
% \begin{tabular}{p{3cm}p{4cm}}
% \hline Free Parameters 	&	 Value \\ 
% \hline \hline 
% $M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
% $\dot{M}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
% $\alpha_X$ 	 &	 $-0.9$ \\ 
% $L_{X} $ 	 &	 $1\times10^{43}~\rm{ergs~s^{-1}}$\\ 
% $r_{disk}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
% $r_{disk}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
% $\dot{M}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
% $r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
% $r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
% $\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
% $\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
% $\lambda$ 	&	 $0$ \\ 
% $v_{\infty}$ 	&	 $v_{esc}$(f=1) \\ 
% $R_v$ 	        &	 $1\times10^{18}$cm \\ 
% $\alpha$ 	&	 $1.0$ \\
% \hline 
% Derived Parameters 	&	 Value \\ 
% \hline \hline
% $L_{\nu}(2500\mbox{\scriptsize{\AA}})$&	 $6.3\times10^{30}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{\nu}(2keV)$   &	 $1.2\times10^{25}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{bol}$ 	 &	 $2.4\times10^{46}~\rm{ergs~s^{-1}}$\\
% $M_{bol}$ 	 &	 -27.4\\ 
% $M_u$            &	 -26.2\\ 
%  $\alpha_{OX} $ 	 &	 -2.2\\ 
% \end{tabular}
% \caption{Wind geometry parameters used in the benchmark model.}
% \label{wind_param}
% \end{table}

@@

Show that clumping stops over-ionization.

@@

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/f2000_m9.png}
\caption
{
$F_\lambda$ at $2000$~\AA\ as a function of inclination from
\agn\ models. Spectra are computed for BH of $10^9~M_\odot$ with
a number of different BH spins and Eddington fractions. The black line
shows a standard multi-temperature blackbody AD model.
}
\label{fig:alpha_ox}
\end{figure}

\section{Summary}

Main points:

\begin{enumerate}
\item We have introduced a simple, first-order treatment 
of clumping in our model, and found that it can now maintain
the required ionization state while agreeing well with the X-ray
properties of AGN/QSOs.
\item We find that clumping also causes a significant 
increase in the strength of the  emission
lines produced by the model. This is true both
of collisionally excited resonance lines (such as \civ, \nv)
and recombination lines (such as \la, \ha\ and the Balmer series).
\item The line EWs in our models are not comparable to those in Quasar composite
spectra. This is due to a fundamental constraint discussed in section ?. If the BLR
emits fairly isotropically then for an anisotropic classical thin accretion disk
it is not possible to achieve line ratios at low inclinations that are comparable to
those at high inclinations. This is a robust conclusion which 
is independent of the assumed BLR geometry. 
\item We have examined the effect of GR on our disk SED, using the disk atmosphere
and GR ray-tracing code AGNSPEC. While including GR effects
does cause the disk SED to become significantly more isotropic,
the effect is not large enough to produce uniform line to continuum ratios
with viewing angle. We discuss other solutions to this problem in section ?; 
It is possible that a combination of GR and reprocessing by the wind could provide a 
solution, and a number of complicated selection effects may be at work
in the building of the quasar composites. Nevertheless,
our conclusions pose a challenge to the current unification picture.
\end{enumerate}

>>>>

\author[Matthews et al.]{J.~H.~Matthews$^1$\thanks{jm8g08@soton.ac.uk}, C.~Knigge$^1$,
N.~Higginbottom$^1$, K.~S.~Long$^2$, S.~A.~Sim$^3$, S.~W.~Mangham$^1$.
\medskip  
\\$^1$School of Physics and Astronomy, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom
\\$^2$Space Telescope Science Institute, 3700 San Martin Drive, Baltimore, MD, 21218
\\$^3$School of Mathematics and Physics, Queens University Belfast, University Road, Belfast, BT7 1NN, Northern Ireland, UK}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%          INTRODUCTION
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

Introduction focussing on key points.
Paragraphs:

\begin{enumerate}
\item standard wind + BALQSO introduction
\item Can an outflow also be the BLR? (unification)
\item some discussion of scales, referencing e.g. reverb maps, variability, microlensing, Arav 
\item AGN variability leading to clumping background: stellar winds, clumping in AGN winds.
\item Modelling efforts so far (Proga, Higginbottom, etc.)
\item Plan of paper
\end{enumerate}
%%\item

\clearpage

% Outflows are ubiquitous in active galactic nuclei (AGN) and quasi-stellar objects (QSOs).
% They can take the form of highly collimated radio jets \citep{bellonijet2010}, 
% or mass-loaded `winds' emanating from the accretion disc. 
% Approximately $20\%$ of QSOs exhibit broad absorption lines (BALs) in the ultraviolet,
% providing clear evidence for outflowing absorbing material
% \citep{weymann1991, knigge2008, turnermiller2009,allen2011}.
% The simplest explanation for the incidence of 
% BALs in quasar samples is in terms of an accretion disc wind (ADW). 
% In this paradigm, the BALQSO fraction is associated with
% the covering factor of the outflow.
% Indeed, many have gone further, and suggested that the diverse phenemonology
% of luminous AGN and QSOs can be broadly explained by
% into a simple picture of geometric unification \citep[e.g.]{MCGV95, elvis2000}. 
% In such a model, a biconical wind rises from 
% a thin accretion disc \cite{shakurasunyaev1973}, approaching velocities
% close to $\sim0.1c$. Depending on viewing angle, an observer 
% may then see a BALQSO, LoBAL-QSO or normal `Type 1' quasar- with 
% broad emission lines superposed on a disc-like continuum.

% The UV spectrum of {\em non-BAL} QSOs is typified by a series
% of strong emission lines (e.g. \la, \civ, \nv) with an underlying blue continuum
% - the so-called {\sl `big blue bump'} (BBB). The BBB is normally attributed to blackbody-like
% emission from an accretion disc surrounding the central black hole (REF), similar
% to that described by \cite{shakurasunyaev1973}. However,
% a number of issues have arisen relating to this model. First, 
% AGN/QSO spectra exhibit a `break' in the spectrum at around $1000$~\AA 
% which scales only weakly with black hole mass. This potentially 
% suggests a problem with a thin disc model (e.g. Antonucci). 
% However, it is possible that this is the result of incorrect intergalactic medium
% (IGM) corrections (REF) or corresponds to the temperature 
% at which a line-driven wind carries mass away from the system (laor davis).
% Second, results from microlensing (REFs) imply that the disc emission 
% region is $\sim4$ times larger than one might expect from a Shakura-Sunyaev
% model. Inhomogenous discs have been proposed as an alternative (REFs).
% These observations appear to pose problems for the thin disc model of luminous,
% sub-Eddington AGN. However, recent results from \cite{capellupo2015} find 
% that if one includes a combination of mass-loss, general relativity (GR) and Comptonisation
% then AGN SEDs can, in general, be fitted well with accretion disc models.
% Uncovering the intrinsic disc SED and understanding the effect of the outflow on the accretion 
% mechanism is clearly crucial if we are to properly understand the physical nature
% of AGN.

% The geometry and size of the BLR is also a matter of contention. 
% The main constraints on the emission region size come from microlensing (REFs)
% and reverberation mapping (REFs). While these observations have mostly been
% carried out for Seyfert galaxies, there are also a few instances of these methods
% being applied to quasars (REFs). A number of different proposals have been made 
% for the origin of the BLR. Early AGN unification scenarios posited that the broad emission lines
% where produced by clouds of plasma orbiting fairly close to the BH (refs).
% Since then, multiple interpretations of a disc wind model have been proposed,
% with varying radii and geometries (refs).

% In this paper, we attempt to test the disc wind paradigm 
% using Monte Carlo radiative transfer and photoionization calculations.
% In section 2, we describe our code. In section 3, we briefly discuss some 
% of the successes and failures of our previous benchmark model for BALQSOs 
% \citep[][hereafter H13]{higginbottom2013} and outline the model, including 
% a description of our clumping prescription. In section 4, we present the results 
% from a clumped model which successfully reproduces
% the observed ionization state while maintaing realistic X-ray properties.
% In section 5 we discuss our results, focussing particular on the anisotropy of 
% disc emission and GR effects, and finally, in section 6, we summarise our findings.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% MCRT

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Ionization and Radiative Transfer}

We use the MCRT code \py\ to carry out our radiative transfer and photoionization
calculations in non-local thermodynamic equilibrium (non-LTE). 
The code is described extensively by a series of authors (LK02, S05, H13, M15).
For that reason we only briefly describe the key elements of the global 
ionization calculation and other important aspects.

\subsection{Line transfer}

To treat line transfer, we adopt the same hybrid scheme 
described by M15...
%in which the energy flows
% through the system are described in terms of indivisible
% energy quanta interacting with either `simple ions'
% or `macro-atoms'. The macro-atom implementation 
% is described in full by \cite{lucy2002, lucy2003}.
% The scheme allows one to treat

@@

Describe the ionization scheme. Combination of H13 and M15 but with rate equations
solved explicitly for ionization solution.

% Macro-atoms have their ion and level populations derived from
% MC rate estimators as described by M15. To calculate the ionization fractions
% of simple-ions the dilute-blackbody modified Saha approach \citep{ML93} is no longer appropriate
% due to the presence of a power-law X-ray source. Instead, we model
% the SED in a cell using the technique described by H13.

% We improve on this method by abandoning the modified Saha approach
% entirely, and instead computing the ionization state by 
% explicitly solving the rate equations in between ions in non-LTE. 
% Photoionization and recombination rates are calculated using MC estimators recorded
% during the photon propagation. Not only does this dispense with a number of
% small assumptions made in the modified Saha approach, it is also more numerically stable, 
% and in principle allows the direct addition of extra physical processes such as Auger ionization,
% which would have to be approximated if using the previous technique.

\subsection{Atomic Data}

% We use the same atomic data as described by LK02 and since
% updated by H13 and M15, with the addition of direct ionization data from Dere (????). 
% Photoionization cross-sections are from Topbase (REF) and 
% \cite{vfky}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% MODEL DESCRIPTION

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bigskip
\bigskip
\bigskip

\section{A Clumpy Biconical Disk Wind Model for QSOs}

Our kinematic prescription for a biconical disc wind model
follows \cite{SV93}, and is described further by
LK02, H13 and M15.

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/fig2_cartoon.eps}
\caption
{
A cartoon showing the geometry and some key parameters of
our biconical wind model.
}
\label{fig:cartoon}
\end{figure}

\subsection{A Simple Biconical Disc Wind Model}

\cite{higginbottom2013} presented a benchmark model for
(BAL)QSOs...introduce key parameters.

\subsubsection{Photon Sources}

Describe the photon sources in the model.

% The accretion disc in our model is geometrically thin, but optically thick
% and this adopt a standard multi-temperature blackbody
% using a \cite{shakurasunyaev1973} temperature profile. 
% The emergent SED is thus determined by the specified accretion rate ($\dot{m}$)
% and central BH mass ($M_{BH}$).
% The inner radius of the disc extends to the innermost stable circular orbit
% (ISCO) of the BH. 
% %% OUTER RADIUS? 
% We assume a Schwarzchild BH with an ISCO at $6~R_G$.
% The X-ray source in our model is treated as an isotropic sphere at the ISCO,
% which emits photons according to a power law,

% \begin{equation}

% \end{equation}

% With a normalisation such that it produces the correct 2-10kev luminosity.

% Photons striking the

% which successfully reproduced BAL features 
% in synthetic spectra using \py. However, the model
% had a few key drawbacks when considering its potential
% for unification. First, it failed to produce significant
% broad emission lines at non-BAL viewing angles; this is 
% a key requirement for a unified QSO model. Second,
% the model was significantly X-ray weak, as it was found
% that an X-ray luminosity comparable to that of a QSO 
% over-ionized the wind, resulting in no UV resonance 
% line absorption features.

% \begin{table}
% \begin{tabular}{p{3cm}p{4cm}}
% \hline Free Parameters 	&	 Value \\ 
% \hline \hline 
% $M_{BH}$ 	 &	 $1\times 10^9~\rm{M_{\odot}}$ \\ 
% $\dot{M}_{acc}$ 	 &	 $5~M_{\odot}yr^{-1} \simeq 0.2~\dot{M}_{Edd}$\\ 
% $\alpha_X$ 	 &	 $-0.9$ \\ 
% $L_{X} $ 	 &	 $1\times10^{43}~\rm{ergs~s^{-1}}$\\ 
% $r_{disc}(min)=r_{X}$   &	 $6r_g=8.8\times10^{14}~{\rm cm}$ \\ 
% $r_{disc}(max)$   &	 $3400r_g = 5\times10^{17}~{\rm cm}$ \\ 
% $\dot{M}_{wind}$  &	 $5~M_{\odot}yr^{-1}$ \\ 
% $r_{min}$ 	&	 $300r_{g} = 4.4\times10^{16}~{\rm cm}$\\ 
% $r_{max}$ 	&	 $600r_{g} = 8.8\times10^{16}~{\rm cm}$ \\ 
% $\theta_{min}$ 	&	 $70.0^{\circ}$ \\ 
% $\theta_{max}$ 	&	 $82.0^{\circ}$ \\ 
% $\lambda$ 	&	 $0$ \\ 
% $v_{\infty}$ 	&	 $v_{esc}$(f=1) \\ 
% $R_v$ 	        &	 $1\times10^{18}$cm \\ 
% $\alpha$ 	&	 $1.0$ \\
% \hline 
% Derived Parameters 	&	 Value \\ 
% \hline \hline
% $L_{\nu}(2500\mbox{\scriptsize{\AA}})$&	 $6.3\times10^{30}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{\nu}(2keV)$   &	 $1.2\times10^{25}~\rm{ergs~s^{-1}~Hz^{-1}}$\\ 
% $L_{bol}$ 	 &	 $2.4\times10^{46}~\rm{ergs~s^{-1}}$\\
% $M_{bol}$ 	 &	 -27.4\\ 
% $M_u$            &	 -26.2\\ 
%  $\alpha_{OX} $ 	 &	 -2.2\\ 
% \end{tabular}
% \caption{Wind geometry parameters used in the benchmark model.}
% \label{wind_param}
% \end{table}

@@

The wind rises slowly from the disc at first, with clumped densities
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity of around $3$ times the escape velocity 
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration means that
the wind exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
By clumping the wind, we are able to produce the range of ionization states observed
in Quasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}1\times10^{43}~\rm{ergs~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

One useful measure of the ionization state is the ionization parameter, $U$, given by
\begin{equation}
U = \frac{4\pi}{n_H c}\int_{13.6{\rm{eV}}}^{\infty}\frac{{J_{\nu}d\nu}}{h\nu}.
\end{equation}
\noindent where $\rm{n_H}$ is the local number density of H, and $\nu$ denotes photon 
frequency. The ionization parameter is a useful measure of the ionization state of a plasma, 
as it evaluates the ratio of the number density of ionizing photons to the local 
H density.

@@

% \begin{figure}
% \centering
% \includegraphics[width=0.45\textwidth]{figures/f2000_m9.eps}
% \caption
% {
% $F_\lambda$ at $2000$~\AA\ as a function of inclination from
% \agn\ models. Spectra are computed for BH of $10^9~M_\odot$ with
% a number of different BH spins and Eddington fractions. The black line
% shows a standard multi-temperature blackbody AD model.
% }
% \label{fig:alpha_ox}
% \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% SUMMARY

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\section{Summary}

We have carried out MCRT simulations using a simple
prescription for a biconical disc wind, with
the aim of expanding on the work of H13 and assessing 
the viability of such a model for geometric unification of quasars.
We find the following main points:

\begin{enumerate}
\item We have introduced a simple, first-order treatment 
of clumping in our model, and found that it can now maintain
the required ionization state while agreeing well with the X-ray
properties of AGN/QSOs.
\smallskip
\item We find that clumping also causes a significant 
increase in the strength of the  emission
lines produced by the model. This is true both
of collisionally excited resonance lines (such as \civ, \nv)
and recombination lines (such as \la, \ha\ and the Balmer series).
\smallskip
\item The line EWs in our models are not comparable to those in Quasar composite
spectra. This is due to a fundamental constraint discussed in section ?. If the BLR
emits fairly isotropically then for a foreshortened, limb-darkened classical thin accretion disc
it is not possible to achieve line ratios at low inclinations that are comparable to
those at high inclinations. This is a robust conclusion which 
is independent of the assumed BLR geometry. 
\smallskip
\item We have examined the effect of GR on our disc SED, using the disc atmosphere
and GR ray-tracing code \agn. While including GR effects
does cause the disc SED to become significantly more isotropic,
the effect is not large enough to produce uniform line to continuum ratios
with viewing angle. We discuss other solutions to this problem in section ?; 
It is possible that a combination of GR and reprocessing by the wind could provide a 
solution, and a number of complicated selection effects may be at work
in the building of the quasar composites.
\end{enumerate}

Our work confirms a number of expected outcomes from such a model, and suggests 
that a simple geometry such as this can come close to explaining much of the 
phenomenology of quasars. Nevertheless, our conclusions pose a clear challenge 
to the current unification picture.

% In particular, we are left with the following possibilities:

% \begin{enumerate}
% \item BALQSOs are not, as a rule, drawn from the quasar population, and 
% instead represent a subclass of objects with strong outflows and differing
% emission line properties to normal luminous AGN and quasars.
% \item A combination of GR effects and reprocessing of the disc continuum
% by the outflow results in a more or less isotropic continuum, explaining
% the comparable line EWs between BALQSOs and quasars. 
% \item The line emission is anisotropic in such a way so as to counter the 
% anisotropy of the disc continuum.
% \end{enumerate}

% We favour a combination of the first two scenarios, based on a combination
% of physical and ontological arguments. It is clear from samples of BALQSOs
% and normal quasars that they have a fairly large range of Eddington fractions, 
% black hole masses and inclinations. The composite spectra will therefore
% be subject to a myriad of complex selection effects.
